---
title: "Stock_Indexes"
author: "Karolina DM"
date: '2022-06-27'
output: html_document
---

---
title: "Stock Indexes"
author: "Karolina DM"
output: html_document
---

```{r setup, include=FALSE}
library("knitr")
library("tidyverse")
library("plotly")
library("htmlwidgets")
source("stock_indexes_analysis.R")
```


## This is a report about stock indexes in the world markets since 24 February 2020 to 17 June 2022 (data collected since the outbreak of the pandemic in Europe until now). For this report I choose selected:

- __Dow Jones Industrial Average__,
- __FTSE China A50__,
- __FTSE 100__,
- __NASDAQ__,
- __NIKKEI 225__,
- __S&P 500__,
- __EURO STOCK 50__,
- __VIX__

## All of data orgin from the website: www.wsj.com/market-data

## For analyzis I performed among others: time series models, multiple regression, cross validation.This methods allows to predict values of stock indexes in the future. The calculations are based on the references at the end of the document.


***

__Dow Jones Industrial Average__  is:

> "a stock index that tracks 30 of the largest U.S. companies. Created in 1896, it is one of the oldest stock indexes, and its performance is widely considered to be a useful indicator of the health of the entire U.S. stock market".

(source of definition: https://www.fool.com/investing/stock-market/indexes/dow-jones/;

source of data: https://www.wsj.com/market-data/quotes/index/DJIA/historical-prices )

__FTSE China A50__  is:

> "an index for 50 stocks of companies with the highest market capitalisation listed on the Shanghai and Shenzhen stock exchanges".

(source of definition: https://www.avatrade.com/trading-info/financial-instruments-index/indices/china-a50;

source of data: https://www.wsj.com/market-data/quotes/index/XX/XIN9/historical-prices )

__FTSE 100__ is:

> "an index composed of the 100 largest (by market capitalisation ) companies listed on the London Stock Exchange (LSE)".

(source of definition: https://www.ii.co.uk/knowledge-centre/quick-guides/before-you-start/what-is-the-ftse-100;

source of data: https://www.wsj.com/market-data/quotes/index/UK/UKX/historical-prices)

__NASDAQ__ is:

> "The first electronic stock market listing over 5000 companies. The Nasdaq stock market comprises two separate markets, namely the Nasdaq National Market, which trades large, active securities and the Nasdaq Smallcap Market that trades emerging growth companies".

(source of definition: https://www.nasdaq.com/glossary/n/nasdaq-stock-market;

source of data: https://www.wsj.com/market-data/quotes/index/NASDAQ/historical-prices)

__NIKKEI 225__ is:

> "the most recognized Japanese stock market index. It comprises Japan’s top 225 companies that are listed on the Tokyo Stock Exchange. The Nikkei Index is considered an important measure of the Japanese stock market and the performance of the Japanese economy."

(source of definition: https://corporatefinanceinstitute.com/resources/knowledge/trading-investing/nikkei-index/;

source of data: https://www.wsj.com/market-data/quotes/index/JP/NIK/historical-prices)

__S&P 500__ is:

> "a market-capitalization-weighted index of 500 leading publicly traded companies in the U.S. It is not an exact list of the top 500 U.S. companies by market cap because there are other criteria that the index includes."

(source of definition: https://www.investopedia.com/terms/s/sp500.asp;

source of data: https://www.wsj.com/market-data/quotes/index/SPX/historical-prices)

__EURO STOCK 50__ :

> "represents the performance of the 50 largest companies among the 20 supersectors in terms of free-float market cap in Eurozone countries. The index has a fixed number of components and is part of the STOXX blue-chip index family. The index captures about 60% of the free-float market cap of the EURO STOXX Total Market Index (TMI)"

(source of definition: https://www.stoxx.com/document/Bookmarks/CurrentFactsheets/SX5GT.pdf;

source of data: https://www.wsj.com/market-data/quotes/index/XX/SX5E/historical-pricess)

__VIX__ is:

> "based on the prices of options on the S&P 500 Index and is calculated by aggregating weighted prices of the index’s call and put options over a wide range of strike prices."

(source of definition: https://corporatefinanceinstitute.com/resources/knowledge/trading-investing/vix-volatility-index/;

source of data: https://www.wsj.com/market-data/quotes/index/VIX/historical-prices)


This plot below presents all indexes:

```{r, dev='png', out.width="130%", echo=FALSE, warning = FALSE}
all_ind_plotly
```

This interactive plot below presents value of EURO STOCK 50:

```{r, dev='png', out.width="130%", echo=FALSE, warning = FALSE}
sx5e
```

A time series is an ordered set of measurements taken at regular intervals, an ideal example of which is the stock exchange indexes. 

Time series plot for EURO STOCK 50 is as follows:

```{r, dev='png', out.width="130%", echo=FALSE, warning = FALSE}
plotSX5Euro
```

Time series wad broke down into: seasonal component, trend, and residuals.

```{r, dev='png', out.width="130%", echo=FALSE, warning = FALSE}
plot(dcmp)
```

We see some seasonal character of our data.

Augmented Dickey-Fuller Test for stationarity confirms that variable EURO STOCK 50 is non-stationary: 

``` {r, echo=FALSE, error=FALSE, warning = FALSE}
adf.test(tsSX5Euro)
```

Then it was done time series evaluation using autocovariance (_acf_ function) and partial autocovariance (_pacf_ function). Autocovariance presents the correlation of the time series with itself shifted by a certain time interval. In turn, partial autovariance is the size of the correlation between the time series and its shift (Lander, 2018, s.404).

This plot shows result of autocovariance:

``` {r, echo=FALSE,  out.width="130%", error=FALSE, warning = FALSE}
acf(tsSX5Euro) 
```

This plot shows result of partial autocovariance:

``` {r, echo=FALSE, out.width="130%", error=FALSE, warning = FALSE}
acf(tsSX5Euro) 
```

The charts confirm the non-stationary nature of the trend. Therefore, a differentiation has to be performed.
The number of differentiations was determined using the _ndiffs_ function and amounted to 0. 


The _arima_ function showed that the optimal model for the discussed time model will be ARMA (1, 0, 0).
The ACF and PACF for the ideal model show the white noise pattern:

``` {r, echo=FALSE, out.width="130%", error=FALSE, warning = FALSE}
acf(best_tsSX5Euro$residuals)
```

``` {r, echo=FALSE, out.width="130%", error=FALSE, warning = FALSE}
pacf(best_tsSX5Euro$residuals)
```

This is the result of building ARIMA model(1,0,0):

``` {r, echo=FALSE, out.width="130%", error=FALSE, warning = FALSE, tidy = TRUE}
summary(fit_tsSX5Euro)
```

This is prediction based on the ARIMA model forecasting for 24 months with the standard error:

``` {r, echo=FALSE, out.width="130%", error=FALSE, warning = FALSE, tidy = TRUE}
plot(theForecastSX5Euro)
```

Prediction was also performed with _Naive_ _Forecasting_ Method. This is a result:

``` {r, echo=FALSE, out.width="130%", error=FALSE, warning = FALSE, tidy = TRUE}
summary(naive_mod)
```

This is a plot: 

``` {r, echo=FALSE, out.width="130%", error=FALSE, warning = FALSE, tidy = TRUE}
naive_plot
```

Prediction was also performed with _Holt's_ _Trend_ Method:

``` {r, echo=FALSE, out.width="130%", error=FALSE, warning = FALSE, tidy = TRUE}
summary(holt_mod)
```

The predictions made by all methods were compared. MAPE and MAE turned out to be the smallest in the __ARIMA model__.

__Multiple regression__ model for EURO STOXX 50 Index is as follows:

``` {r, echo=FALSE, out.width="130%", error=FALSE, warning = FALSE, tidy = TRUE}
summary(model1)
```

The model explains 96% of the variability of EURO STOXX 50 Index, and all the variables used in the model are significant.

``` {r, echo=FALSE, out.width="130%", error=FALSE, warning = FALSE, tidy = TRUE}

ols_test_normality(model1) 
ols_plot_resid_hist(model1) 

```

The residuals in model meet the assumptions of the normal distribution, and the result of the K-S test also it confirms (p > 0,05 that is, there is no reason to reject the null hypothesis and the distribution is normal). 

``` {r, echo=FALSE, out.width="130%", error=FALSE, warning = FALSE, tidy = TRUE}
ols_plot_resid_qq(model1) 
```

The Quantile-Quantile plot also confirms the normal distribution of EURO STOXX 50 Index -
the observations are almost perfectly positioned on the straight line (except for the rest).

``` {r, echo=FALSE, out.width="130%", error=FALSE, warning = FALSE, tidy = TRUE}
cook
```

The _Cook's_ plot above allows you to explore outliers.

Multiple regression models were also compared. The following regression models were built:

* model1 <- lm(SX5Euro_Close ~ DJ_Close + FTSE_Close + FTSE100_Close + NASDAQ_Close + 
              NIKKEI_Close + SP500_Close + VIX_Close, data = all_indexes)
* model2 <-  lm(SX5Euro_Close ~ DJ_Close +  FTSE100_Close + NASDAQ_Close + 
            SP500_Close + VIX_Close, data = all_indexes)
* model3 <-  lm(SX5Euro_Close ~ VIX_Close + NASDAQ_Close + SP500_Close, data = all_indexes)
* model4 <-  lm(SX5Euro_Close ~ VIX_Close + FTSE100_Close, data = all_indexes)
* model5 <- lm(SX5Euro_Close ~ NIKKEI_Close + FTSE_Close, data = all_indexes)
* model6 <- lm(SX5Euro_Close ~ DJ_Close, data = all_indexes)

They were visualized using the multiplot function.

``` {r, echo=FALSE, out.width="130%", error=FALSE, warning = FALSE, tidy = TRUE}
multiplot_models
```


``` {r, echo=FALSE,  error=FALSE, warning = FALSE, tidy = TRUE}
anova(model1nona, model2nona, model3nona, model4nona, model5nona, model6nona)
```

The smallest RSS is observed in the case of model 1. This is also confirmed by the results of AIC ( _Akaike_ _Information_ _Criterion_): __`r AIC(model1nona, model2nona, model3nona, model4nona, model5nona, model6nona)`__
and BIC ( _Bayesian_ _Information_ _Criterion_): __`r BIC(model1, model2, model3, model4, model5, model6)`__.

Due to the obsolescence of using ANOVA to test regression models (Lander, 2018, p. 337), a cross-validation with generalized linear models was also performed.

``` {r, echo=FALSE,  error=FALSE, warning = FALSE, tidy = TRUE}
models_results
```

It has been shown once again that the first model is characterized by the lowest error value.

***

In summary, we see fertility rate and birth rate falling over the years, while death rate and life expectancy are increasing. It is very important to follow demographic changes in countries and individual parts of the world so that governments can react early and shape their policies on public health, pro-family policies, etc.It is impossible to track demographic changes in isolation from the data on armed conflicts and the related migration of people, climate change, and the pandemic situation. 

***
References:

* Biecek, P., Trajkowski, K. (2019). Na przełaj przez Data Mining z pakietem R. Retrived from:  https://pbiecek.github.io/NaPrzelajDataMiningR/part-7.html (05.07.2022).
* Eulogio, R. (2018). Performing a Time-Series Analysis on the S&P 500 Stock Index. 
Retrived from:  https://blogs.oracle.com/ai-and-datascience/post/performing-a-time-series-analysis-on-the-sampp-500-stock-index (30.06.2022).
* Long, D. (2019). R Cookbook, 2nd Edition. Retrived from: https://rc2e.com (05.07.2022).
* Lander, J. P. (2018). R dla każdego. Zaawansowane analizy i grafika statystyczna. Warszawa: APN Promise.
* Singh, D. (2019). Time seres forecasting using R. Retrived from: https://www.pluralsight.com/guides/time-series-forecasting-using-r (30.06.2022).
* Statology (2021). Retrived from: https://www.statology.org/dickey-fuller-test-in-r/ (01.07.2022).
* Wawrowski, R. Metody przetwarzania i analizy danych w R. Retrieved from: http://www.wawrowski.edu.pl/adr/index.html (29.06.2022).



