---
title: "Stock_Indexes"
author: "Karolina DM"
date: '2022-06-27'
output: html_document
---

---
title: "Stock Indexes"
author: "Karolina DM"
output: html_document
---

```{r setup, include=FALSE}
library("knitr")
library("tidyverse")
library("plotly")
library("htmlwidgets")
source("stock_indexes_analysis.R")
```


## This is a report about stock indexes in the world markets since February 2020 to July 2022. For this report I choose selected:

- __Dow Jones Industrial Average__,
- __FTSE China A50__,
- __FTSE 100__,
- __NASDAQ__,
- __NIKKEI 225__,
- __S&P 500__,
- __EURO STOCK 50__,
- __VIX__


***

__Dow Jones Industrial Average__  is:

> "a stock index that tracks 30 of the largest U.S. companies. Created in 1896, it is one of the oldest stock indexes, and its performance is widely considered to be a useful indicator of the health of the entire U.S. stock market".

(source of definition: https://www.fool.com/investing/stock-market/indexes/dow-jones/;
source of data: https://www.wsj.com/market-data/quotes/index/DJIA/historical-prices )

__FTSE China A50__  is:

> "a stock index that tracks 30 of the largest U.S. companies".

(source of definition: https://www.fool.com/investing/stock-market/indexes/dow-jones/;
source of data: https://www.wsj.com/market-data/quotes/index/DJIA/historical-prices )


This plot presents all indexes:

```{r plot_all=plot, dev='png', out.width="150%", echo=FALSE, warning = FALSE}
plot(plot_all)
```









***

In summary, we see fertility rate and birth rate falling over the years, while death rate and life expectancy are increasing. It is very important to follow demographic changes in countries and individual parts of the world so that governments can react early and shape their policies on public health, pro-family policies, etc.It is impossible to track demographic changes in isolation from the data on armed conflicts and the related migration of people, climate change, and the pandemic situation. 